---
title: "Model Comparison - Postpartum"
author: "isken"
date: "June 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(stringr)
library(dplyr)
#library(Metrics)
library(caret)
library(MASS)
```

```{r}
source("lib/caret_ptps.r")
```


## Introduction

In this second set of comparisons, let's focus on LDR occupancy in terms of output variables. This 
should be more difficult than postpartum predictions since occupancy of LDR is impacted by time queueing in obs as well as time blocked in LDR.


The output variables to predict will be mean occupancy, 75th and 95th percentiles of occupancy. 

## Experimental design

All 150 simulated scenarios will be used. 

The models to be compared are:

- load and square root based simple queueing approximation based linear models
- multiple linear regression
- k-nearest neighbor
- regression trees
- cubic splines
- neural networks

Comparisons will be made using 5-crossfold validation with RMSE as the error metric. In addition, we'll create a holdout test set of 1/6 (25 scenarios) in which a final RMSE comparison will be done.

## Read input/output data

```{r}
obnetwork_df <- read.csv("mmdata/obnetwork_approx_vs_sim.csv")

```
Load the simulation inputs and results

```{r}
load("mmdata/train_exp9_tandem05_nodischadj.RData")
```

Merge the inputs and results

```{r}
simresults_obnetwork <- merge(results_df, obnetwork_df, by="scenario")
```


Select relevant columns. For postpartum comparisons, prior knowledge suggests that only the postpartum related input variables be used. We will assume for now that enough "process physics" knowledge is present that the analyst would create the engineered features `load_ldr` <- lam_ldr * alos_ldr` and `rho_ldr <- load_ldr/cap_ldr`. It would be interesting to also run a set of experiments in which we relax this assumption (a naive analyst scenario).

```{r}
modeling_cols <- c('lam_obs', 'alos_obs', 'cap_obs', 'load_obs', 'rho_obs',
                   'lam_ldr', 'alos_ldr', 'cap_ldr', 'load_ldr', 'rho_ldr',
                   'lam_pp', 'alos_pp', 'cap_pp', 'load_pp', 'rho_pp',
                   'cv2_obs', 'cv2_ldr', 'cv2_ldr', 'tot_c_rate',
                   'prob_blockedby_pp_approx',
                   'occ_mean_mean_ldr', 'occ_mean_p75_ldr', 'occ_mean_p95_ldr',
                   'mean_pct_blocked_by_pp', 'mean_blocked_by_pp_mean', 
                   'mean_blocked_by_pp_p95')

modeling_df <- simresults_obnetwork[,modeling_cols]
```

## Input args

Build up the formula argument.

```{r}
model_formula_lhs <- 'occ_mean_mean_ldr'
model_formula_rhs_ldr <- 'cap_ldr + load_ldr + rho_ldr'
model_formula_rhs_ldr2 <- 'I(cap_ldr^2) + I(load_ldr^2) + I(rho_ldr^2)'
model_formula_rhs_pp <- 'alos_pp + cap_pp + load_pp + rho_pp'
model_formula_rhs_pp2 <- 'I(alos_pp^2) + I(cap_pp^2) + I(load_pp^2) + I(rho_pp^2)'

model_formula_rhs <- str_c(model_formula_rhs_ldr, 
                           model_formula_rhs_ldr2, 
                           model_formula_rhs_pp,
                           model_formula_rhs_pp2,
                           sep = " + ")

model_formula <- as.formula(str_c(model_formula_lhs, model_formula_rhs, sep = " ~ "))
```

Rest of input args for caret_ptps

```{r}
pm_unit_model_id <- 'occmean_ldr_poly1'

method <- 'lm'
perf_measure <- 'occmean'
unit <- 'ldr'
model <- 'lmpoly1'
pct_train_val <- 0.7
seed.partition <- 37
```

```{r}
test_lm1 <- lm(model_formula, data = modeling_df)
summary(test_lm1)
```

```{r}
summary(modeling_df$load_ldr)
```


```{r}
ptps.model_results <- caret_ptps(pm_unit_model_id = pm_unit_model_id, 
                                 model_formula = model_formula, 
                                 data = modeling_df, 
                                 method = method,
                                 perf_measure = perf_measure,
                                 unit = unit, 
                                 model = model,
                                 pct_train_val = pct_train_val,
                                 seed.partition = 37,
                                 seed.resample = 0)
```


```{r}
set.seed(train_test_seed)
trainrecs_caret <- createDataPartition(simresults_obnetwork$occ_mean_mean_ldr, 
                                       p = train_proportion, 
                                       list = FALSE, times = 1)

modeling_test_caret_df <- modeling_df[-trainrecs_caret, ]
modeling_train_caret_df <- modeling_df[trainrecs_caret, ]
```

## Control training process

Specify how the model validation will work.

```{r}
fitControl <- trainControl(## 5-fold CV
                           method = "repeatedcv",
                           number = 5,
                           ## repeated ten times
                           repeats = 10)
```

Now we can use the `caret_tps()` function we wrote to quickly evaluate a bunch of models for several
response variables.

```{r}
resample_seed <- 67
```


```{r meanmodels1}
formula_occmean_ldr_load1 <- occ_mean_mean_ldr ~ 
                                  rho_ldr * load_ldr + 
                                  rho_ldr * I(load_ldr ^ 0.5) + 
                                  rho_pp * load_pp + rho_pp * I(load_pp ^ 0.5) 

set.seed(resample_seed)
results_occmean_ldr_load1 <- caret_tps(formula_occmean_ldr_load1, modeling_train_caret_df, 
                                       modeling_test_caret_df,
                                      method="lm", control=fitControl,
                                      scenario="occmean_ldr_load1")

formula_occmean_ldr_load2 <- occ_mean_mean_ldr ~ 
                                  rho_ldr * load_ldr + 
                                  rho_ldr * I(load_ldr ^ 0.5) + 
                                  rho_pp * load_pp + rho_pp * I(load_pp ^ 0.5) + tot_c_rate

set.seed(resample_seed)
results_occmean_ldr_load2 <- caret_tps(formula_occmean_ldr_load2, modeling_train_caret_df, 
                                       modeling_test_caret_df,
                                      method="lm", control=fitControl,
                                      scenario="occmean_ldr_load2")
```

We can use the following base formula for several different techniques.


```{r baseformula}
formula_occmean_ldr_base <- occ_mean_mean_ldr ~ 
                               lam_ldr + alos_ldr + cap_ldr + load_ldr + rho_ldr + 
                               lam_pp + alos_pp + cap_pp + load_pp + rho_pp +
                               tot_c_rate
```

Linear stepwise, cubic splines and a random forest can all use the same basic setup. The only change
is the `method` parameter passed to `caret_tps`. For a basic neural network we have to add the
the `linout=TRUE` parameter so that regression output is given instead of classification output. For
kNN, we have to add a preprocessing parameter to center and scale the inputs. Both of these things
are easy to do since `caret_tps()` can take additional parameters via `...`.

```{r meanmodels2}
set.seed(resample_seed)
results_occmean_ldr_lmstep1 <- caret_tps(formula_occmean_ldr_base, modeling_train_caret_df, 
                                         modeling_test_caret_df,
                                         method="lmStepAIC", control=fitControl,
                                         scenario="occmean_ldr_lmstep1")

set.seed(resample_seed)
results_occmean_ldr_nnet1 <- caret_tps(formula_occmean_ldr_base, modeling_train_caret_df, 
                                       modeling_test_caret_df,
                                       method="nnet", control=fitControl,
                                       scenario="occmean_ldr_nnet1", linout=TRUE)

set.seed(resample_seed)
results_occmean_ldr_rf1 <- caret_tps(formula_occmean_ldr_base, modeling_train_caret_df, 
                                     modeling_test_caret_df,
                                     method="rf", control=fitControl,
                                     scenario="occmean_ldr_rf1")

set.seed(resample_seed)
results_occmean_ldr_knn1 <- caret_tps(formula_occmean_ldr_base, modeling_train_caret_df, 
                                      modeling_test_caret_df,
                                      method="knn", control=fitControl,
                                      scenario="occmean_ldr_knn1", 
                                      preProcess=c("center", "scale"))

set.seed(resample_seed)
results_occmean_ldr_spline1 <- caret_tps(formula_occmean_ldr_base, modeling_train_caret_df, 
                                    modeling_test_caret_df,
                                    method="gamSpline", control=fitControl, 
                                    scenario="occmean_ldr_spline1")
```

```{r rmse_occmean}
cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_load1$scenario, results_occmean_ldr_load1$rmse_train, results_occmean_ldr_load1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_load2$scenario, results_occmean_ldr_load2$rmse_train, results_occmean_ldr_load2$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_lmstep1$scenario, results_occmean_ldr_lmstep1$rmse_train, results_occmean_ldr_lmstep1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_nnet1$scenario, results_occmean_ldr_nnet1$rmse_train, results_occmean_ldr_nnet1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_rf1$scenario, results_occmean_ldr_rf1$rmse_train, results_occmean_ldr_rf1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_knn1$scenario, results_occmean_ldr_knn1$rmse_train, results_occmean_ldr_knn1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occmean_ldr_spline1$scenario, results_occmean_ldr_spline1$rmse_train, results_occmean_ldr_spline1$rmse_test))
```

```{r resamp_mean}
resamps_occmean <- resamples(list(load1 = results_occmean_ldr_load1$caretFit,
                                  load2 = results_occmean_ldr_load2$caretFit,
                          lmstep1 = results_occmean_ldr_lmstep1$caretFit,
                          knn1 = results_occmean_ldr_knn1$caretFit,
                          nnet1 = results_occmean_ldr_nnet1$caretFit,
                          rf1 = results_occmean_ldr_rf1$caretFit,
                          spline1 = results_occmean_ldr_spline1$caretFit))

resamps_occmean
summary(resamps_occmean)
bwplot(resamps_occmean, scales = list(relation = "free"))
dotplot(resamps_occmean, scales = list(relation = "free"))
```


## 95th percentile

```{r}
set.seed(train_test_seed)
trainrecs_caret <- createDataPartition(simresults_obnetwork$occ_mean_p95_ldr, p = train_proportion, 
                                  list = FALSE, times = 1)

modeling_test_caret_df <- simresults_obnetwork[-trainrecs_caret, ]
modeling_train_caret_df <- simresults_obnetwork[trainrecs_caret, ]
```



```{r occ95models1}
formula_occp95_ldr_load1 <- occ_mean_p95_ldr ~ 
                               rho_ldr * load_ldr + rho_ldr * I(load_ldr ^ 0.5) + 
                               rho_pp * load_pp + rho_pp * I(load_pp ^ 0.5) 

set.seed(resample_seed)
results_occp95_ldr_load1 <- caret_tps(formula_occp95_ldr_load1, modeling_train_caret_df, 
                                      modeling_test_caret_df,
                                      method="lm", control=fitControl,
                                      scenario="occp95_ldr_load1")

formula_occp95_ldr_load2 <- occ_mean_p95_ldr ~ 
                               rho_ldr * load_ldr + rho_ldr * I(load_ldr ^ 0.5) + 
                               rho_pp * load_pp + rho_pp * I(load_pp ^ 0.5) + tot_c_rate

set.seed(resample_seed)
results_occp95_ldr_load2 <- caret_tps(formula_occp95_ldr_load2, modeling_train_caret_df, 
                                      modeling_test_caret_df,
                                      method="lm", control=fitControl,
                                      scenario="occp95_ldr_load2")
```

```{r occ95formulabase}
formula_occp95_ldr_base <- occ_mean_p95_ldr ~ 
                              lam_ldr + alos_ldr + cap_ldr + load_ldr + rho_ldr + 
                              lam_pp + alos_pp + cap_pp + load_pp + rho_pp + tot_c_rate
```

```{r occ95models2}
set.seed(resample_seed)
results_occp95_ldr_lmstep1 <- caret_tps(formula_occp95_ldr_base, modeling_train_caret_df, 
                                        modeling_test_caret_df,
                                        method="lmStepAIC", control=fitControl, 
                                        scenario="occp95_ldr_lmstep1")

set.seed(resample_seed)
results_occp95_ldr_nnet1 <- caret_tps(formula_occp95_ldr_base, modeling_train_caret_df, 
                                      modeling_test_caret_df,
                                      method="nnet", control=fitControl, 
                                      scenario="occp95_ldr_nnet1", linout=TRUE)

set.seed(resample_seed)
results_occp95_ldr_rf1 <- caret_tps(formula_occp95_ldr_base, modeling_train_caret_df, 
                                    modeling_test_caret_df,
                                    method="rf", control=fitControl, scenario="occp95_ldr_rf1")

set.seed(resample_seed)
results_occp95_ldr_knn1 <- caret_tps(formula_occp95_ldr_base, modeling_train_caret_df, 
                                     modeling_test_caret_df,
                                     method="knn", control=fitControl,
                                     scenario="occp95_ldr_knn1", 
                                     preProcess=c("center", "scale"))

set.seed(resample_seed)
results_occp95_ldr_spline1 <- caret_tps(formula_occp95_ldr_base, modeling_train_caret_df, 
                                    modeling_test_caret_df,
                                    method="gamSpline", control=fitControl, 
                                    scenario="occp95_ldr_spline1")

```

```{r rmse_occp95}
cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_load1$scenario, results_occp95_ldr_load1$rmse_train, results_occp95_ldr_load1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_load2$scenario, results_occp95_ldr_load2$rmse_train, results_occp95_ldr_load2$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_lmstep1$scenario, results_occp95_ldr_lmstep1$rmse_train, results_occp95_ldr_lmstep1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_nnet1$scenario, results_occp95_ldr_nnet1$rmse_train, results_occp95_ldr_nnet1$rmse_test))

cat(sprintf("%s:\t\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_rf1$scenario, results_occp95_ldr_rf1$rmse_train, results_occp95_ldr_rf1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_knn1$scenario, results_occp95_ldr_knn1$rmse_train, results_occp95_ldr_knn1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_occp95_ldr_spline1$scenario, results_occp95_ldr_spline1$rmse_train, results_occp95_ldr_spline1$rmse_test))
```

```{r}
summary(results_occp95_ldr_spline1$caretFit$finalModel)
```

```{r}
results_occp95_ldr_load1$g_act_vs_pred
results_occp95_ldr_nnet1$g_act_vs_pred
results_occp95_ldr_spline1$g_act_vs_pred
```

```{r resamp_occp95}
resamps_occp95 <- resamples(list(load1 = results_occp95_ldr_load1$caretFit,
                                 load2 = results_occp95_ldr_load2$caretFit,
                          lmstep1 = results_occp95_ldr_lmstep1$caretFit,
                          knn1 = results_occp95_ldr_knn1$caretFit,
                          nnet1 = results_occp95_ldr_nnet1$caretFit,
                          rf1 = results_occp95_ldr_rf1$caretFit,
                          spline1 = results_occp95_ldr_spline1$caretFit))

resamps_occp95
summary(resamps_occp95)
bwplot(resamps_occp95, scales = list(relation = "free"))
dotplot(resamps_occp95, scales = list(relation = "free"))
```

## Repeat above for prob of being blocked by PP

```{r}
set.seed(train_test_seed)
trainrecs_caret <- createDataPartition(simresults_obnetwork$mean_pct_blocked_by_pp, p = train_proportion, 
                                  list = FALSE, times = 1)

modeling_test_caret_df <- simresults_obnetwork[-trainrecs_caret, ]
modeling_train_caret_df <- simresults_obnetwork[trainrecs_caret, ]
```

```{r}
formula_pctblkpp_ldr_mgc1 <- mean_pct_blocked_by_pp ~ prob_blockedby_pp_approx

set.seed(resample_seed)
results_pctblkpp_ldr_mgc1 <- caret_tps(formula_pctblkpp_ldr_mgc1, 
                                       modeling_train_caret_df, modeling_test_caret_df,
                                       method="lm", control=fitControl, 
                                       scenario="pctblkpp_ldr_load1")
```

```{r}
formula_pctblkpp_ldr_base <- mean_pct_blocked_by_pp ~ lam_ldr + alos_ldr + cap_ldr + 
                                           load_ldr + rho_ldr + 
                                           lam_pp + alos_pp + cap_pp + load_pp + rho_pp + tot_c_rate
```

```{r}
set.seed(resample_seed)
results_pctblkpp_ldr_lmstep1 <- caret_tps(formula_pctblkpp_ldr_base, modeling_train_caret_df, 
                                          modeling_test_caret_df,
                                          method="lmStepAIC", control=fitControl,
                                          scenario="pctblkpp_ldr_lmstep1")

set.seed(resample_seed)
results_pctblkpp_ldr_nnet1 <- caret_tps(formula_pctblkpp_ldr_base, modeling_train_caret_df,
                                        modeling_test_caret_df,
                                        method="nnet", control=fitControl,
                                        scenario="pctblkpp_ldr_nnet1", linout=TRUE)

set.seed(resample_seed)
results_pctblkpp_ldr_rf1 <- caret_tps(formula_pctblkpp_ldr_base, modeling_train_caret_df,
                                      modeling_test_caret_df,
                                      method="rf", control=fitControl,
                                      scenario="pctblkpp_ldr_rf1")

set.seed(resample_seed)
results_pctblkpp_ldr_knn1 <- caret_tps(formula_pctblkpp_ldr_base, modeling_train_caret_df,
                                       modeling_test_caret_df,
                                       method="knn", control=fitControl,
                                       scenario="pctblkpp_ldr_knn1", 
                                       preProcess=c("center", "scale"))

set.seed(resample_seed)
results_pctblkpp_ldr_spline1 <- caret_tps(formula_pctblkpp_ldr_base, modeling_train_caret_df, 
                                    modeling_test_caret_df,
                                    method="gamSpline", control=fitControl, 
                                    scenario="pctblkpp_ldr_spline1")
```

```{r rmse_pctblkpp_ldr}
cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_pctblkpp_ldr_mgc1$scenario, results_pctblkpp_ldr_mgc1$rmse_train, results_pctblkpp_ldr_mgc1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_pctblkpp_ldr_lmstep1$scenario, results_pctblkpp_ldr_lmstep1$rmse_train, results_pctblkpp_ldr_lmstep1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_pctblkpp_ldr_nnet1$scenario, results_pctblkpp_ldr_nnet1$rmse_train, results_pctblkpp_ldr_nnet1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_pctblkpp_ldr_rf1$scenario, results_pctblkpp_ldr_rf1$rmse_train, results_pctblkpp_ldr_rf1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_pctblkpp_ldr_knn1$scenario, results_pctblkpp_ldr_knn1$rmse_train, results_pctblkpp_ldr_knn1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_pctblkpp_ldr_spline1$scenario, results_pctblkpp_ldr_spline1$rmse_train, results_pctblkpp_ldr_spline1$rmse_test))
```

```{r resamp_pctblkpp}
resamps_pctblkpp <- resamples(list(mgc1 = results_pctblkpp_ldr_mgc1$caretFit,
                          lmstep1 = results_pctblkpp_ldr_lmstep1$caretFit,
                          knn1 = results_pctblkpp_ldr_knn1$caretFit,
                          nnet1 = results_pctblkpp_ldr_nnet1$caretFit,
                          rf1 = results_pctblkpp_ldr_rf1$caretFit,
                          spline1 = results_pctblkpp_ldr_spline1$caretFit))

resamps_pctblkpp
summary(resamps_pctblkpp)
bwplot(resamps_pctblkpp, scales = list(relation = "free"))
dotplot(resamps_pctblkpp, scales = list(relation = "free"))
```

Note how well mgc does across samples. Expected since it takes physics into account.

## Repeat above for cond mean time blocked by pp

```{r}
set.seed(train_test_seed)
trainrecs_caret <- createDataPartition(simresults_obnetwork$mean_blocked_by_pp_mean, p = train_proportion, 
                                  list = FALSE, times = 1)

modeling_test_caret_df <- simresults_obnetwork[-trainrecs_caret, ]
modeling_train_caret_df <- simresults_obnetwork[trainrecs_caret, ]
```

```{r}
formula_meanblkpp_ldr_mgc2 <- mean_blocked_by_pp_mean ~ condmeantime_blockedbypp_approx

set.seed(resample_seed)
results_meanblkpp_ldr_mgc2 <- caret_tps(formula_meanblkpp_ldr_mgc2, 
                                       modeling_train_caret_df, modeling_test_caret_df,
                                       method="lm", control=fitControl, 
                                       scenario="meanblkpp_ldr_load1")
```

```{r}
formula_meanblkpp_ldr_base <- mean_blocked_by_pp_mean ~ lam_ldr + alos_ldr + cap_ldr + 
                                           load_ldr + rho_ldr + 
                                           lam_pp + alos_pp + cap_pp + load_pp + rho_pp + tot_c_rate
```

```{r}
set.seed(resample_seed)
results_meanblkpp_ldr_lmstep1 <- caret_tps(formula_meanblkpp_ldr_base, modeling_train_caret_df, 
                                           modeling_test_caret_df,
                                           method="lmStepAIC", control=fitControl, 
                                           scenario="meanblkpp_ldr_lmstep1")

set.seed(resample_seed)
results_meanblkpp_ldr_nnet1 <- caret_tps(formula_meanblkpp_ldr_base, modeling_train_caret_df, 
                                         modeling_test_caret_df,
                                         method="nnet", control=fitControl, 
                                         scenario="meanblkpp_ldr_nnet1", linout=TRUE)

set.seed(resample_seed)
results_meanblkpp_ldr_rf1 <- caret_tps(formula_meanblkpp_ldr_base, modeling_train_caret_df, 
                                       modeling_test_caret_df,
                                       method="rf", control=fitControl, 
                                       scenario="meanblkpp_ldr_rf1")

set.seed(resample_seed)
results_meanblkpp_ldr_knn1 <- caret_tps(formula_meanblkpp_ldr_base, modeling_train_caret_df, 
                                        modeling_test_caret_df,
                                        method="knn", control=fitControl, 
                                        scenario="meanblkpp_ldr_knn1", 
                                        preProcess=c("center", "scale"))

set.seed(resample_seed)
results_meanblkpp_ldr_spline1 <- caret_tps(formula_meanblkpp_ldr_base, modeling_train_caret_df, 
                                    modeling_test_caret_df,
                                    method="gamSpline", control=fitControl, 
                                    scenario="meanblkpp_ldr_spline1")
```

```{r rmse_meanblkpp_ldr}
cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_meanblkpp_ldr_mgc2$scenario, results_meanblkpp_ldr_mgc2$rmse_train, results_meanblkpp_ldr_mgc2$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_meanblkpp_ldr_lmstep1$scenario, results_meanblkpp_ldr_lmstep1$rmse_train, results_meanblkpp_ldr_lmstep1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_meanblkpp_ldr_nnet1$scenario, results_meanblkpp_ldr_nnet1$rmse_train, results_meanblkpp_ldr_nnet1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_meanblkpp_ldr_rf1$scenario, results_meanblkpp_ldr_rf1$rmse_train, results_meanblkpp_ldr_rf1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_meanblkpp_ldr_knn1$scenario, results_meanblkpp_ldr_knn1$rmse_train, results_meanblkpp_ldr_knn1$rmse_test))

cat(sprintf("%s:\t rmse_train = %0.5f rmse_test = %0.5f \n", results_meanblkpp_ldr_spline1$scenario, results_meanblkpp_ldr_spline1$rmse_train, results_meanblkpp_ldr_spline1$rmse_test))
```

```{r}
resamps_meanblkpp <- resamples(list(mgc2 = results_meanblkpp_ldr_mgc2$caretFit,
                          lmstep1 = results_meanblkpp_ldr_lmstep1$caretFit,
                          knn1 = results_meanblkpp_ldr_knn1$caretFit,
                          nnet1 = results_meanblkpp_ldr_nnet1$caretFit,
                          rf1 = results_meanblkpp_ldr_rf1$caretFit,
                          spline1 = results_meanblkpp_ldr_spline1$caretFit))

resamps_meanblkpp
summary(resamps_meanblkpp)
bwplot(resamps_meanblkpp, scales = list(relation = "free"))
dotplot(resamps_meanblkpp, scales = list(relation = "free"))
```
So, ML workhorses like neural nets, splines and random forests are tough to beat without good domain specific approximations. A few things to explore:

* what if load and rho not included for the naive ML case?
* how are the results above impacted by validation set size and by number of validation sets for which the whole process is repeated? How much variation in estimated test error?
* how do the techniques compare when extrapolating beyond (or within but not on) data used to build the models?
How stable are the models? How does one measure stability for things like random forests or neural nets?
* is RMSE the "right" error metric to be using?
* does the number of folds in our k-cross matter?
* how amenable are the various techniques to embedding in capacity planning models?
* are results of standard capacity planning techniques impacted by which of the above models is used?
* what about logistic form for prob of block?
* why knn weird training in PP? Does it have something to do with the centering and scaling done during preprocessing?
